---
title: "Measuring Bias and Variance - Regression"
output: html_document
---

```{r}
#Reading the dataset
rm(list = ls(all=TRUE))
data = read.csv('airfoil_self_noise.csv', header = TRUE)
```
# Exploratory Data Analysis
```{r}
head(data)
```

```{r}
tail(data)
```

```{r}
summary(data)
```

```{r}
str(data)
```

```{r}
colSums(is.na(data))
```
#NO missing Values Found
```{r}
colnames(data)
```

```{r}
dim(data)
```
# Splitting the data into train and test
```{r}
library(caret)
set.seed(123)
train_rows <- createDataPartition(data$scaled_sound_pressure_level,p = 0.7,list = FALSE)
traindf = data[train_rows,]
testdf = data[-train_rows,]
```

```{r}
str(testdf)
```

-----------------------------------------------Linear Regression---------------------------------------------
# Bootstrap Sampling and Modeling(100samples, 100models)
```{r}
# parallel   ---------------------------------------------------------
library(foreach)
library(doParallel)

#setup parallel back end to use 8 processors
cl<-makeCluster(6)
registerDoParallel(cl)

#  sample data 100 times 
length_divisor <- 1
predictions<-foreach(m=1:100,.combine=cbind) %dopar% { 
        # using sample function without seed
     sampleRows <- sample(nrow(traindf), size=floor((nrow(traindf)/length_divisor)),replace = TRUE)
     fit <- lm(scaled_sound_pressure_level ~ ., data = traindf[sampleRows,])
     predictions <- data.frame(predict(object=fit, testdf[,-6], se.fit = TRUE)[[1]])
} 
stopCluster(cl)

library(pROC)
dim(predictions)
test_with_predictions = cbind(testdf,predictions)
dim(test_with_predictions)
tail(tail(test_with_predictions))
test_with_predictions = subset(test_with_predictions[,c(-1,-2,-3,-4,-5)])
tail(test_with_predictions)
# Calculating the expected estimate
test_with_predictions$Avg_prediction = rowMeans(test_with_predictions[,2:101])
test_with_predictions$Avg_prediction
tail(test_with_predictions)
```

```{r}
predictions
```

```{r}
dim(predictions)
```
#Calculating Bias Square and Variance
```{r}
library(matrixStats)
test_with_predictions$Bias_square = (test_with_predictions$scaled_sound_pressure_level - test_with_predictions$Avg_prediction)^2
Linear_Regression_Bias_Square = mean(test_with_predictions$Bias_square)
test_with_predictions$Variance = rowVars(x = data.matrix(test_with_predictions, rownames.force = NA), cols = c(-1,-102,-103))
tail(test_with_predictions)
```
#Caculating the model Bias Square
```{r}
Bias_square_LR = mean(test_with_predictions$Bias_square)
Bias_square_LR
```
#Calculating the model Variance
```{r}
Variance_LR = mean(test_with_predictions$Variance)
Variance_LR
```

```{r}
Bias_square_LR + Variance_LR
```
# calculating the Mean Squared Error
```{r}
calculationLR = test_with_predictions[1:101]
calculationLR[2:101] = (calculationLR[2:101]-calculationLR[,1])^2
tail(calculationLR)
 
calculationLR$MSE = rowMeans(calculationLR[,2:101])
mean(calculationLR$MSE)


```
# Calculating MSE
```{r}
calculationLR$MSE = rowMeans(calculationLR[,2:101])
mean(calculationLR$MSE)
```
----------------------------------------------Decision_Tree--------------------------------------------------
# Bootstrap Sampling and Modeling(100samples, 100models)
```{r}
# parallel   ---------------------------------------------------------
library(foreach)
library(doParallel)
library(rpart)

#setup parallel back end to use 8 processors
cl<-makeCluster(6)
registerDoParallel(cl)

#  sample data 100 times 
length_divisor <- 1
predictions_dt<-foreach(m=1:100,.combine=cbind) %dopar% { 
        # using sample function without seed
  library(rpart)   
  sampleRows <- sample(nrow(traindf), size=floor((nrow(traindf)/length_divisor)),replace = TRUE)
     fit_dt <- rpart(scaled_sound_pressure_level ~ .,method="anova", data = traindf[sampleRows,])
     predictions_dt <- data.frame(predict(object=fit_dt, testdf[,-6], se.fit = TRUE))
} 
stopCluster(cl)

library(pROC)
# Calculating the expected estimate
dim(predictions_dt)
test_with_predictions_dt = cbind(testdf,predictions_dt)
dim(test_with_predictions_dt)
tail(tail(test_with_predictions_dt))
test_with_predictions_dt = subset(test_with_predictions_dt[,c(-1,-2,-3,-4,-5)])
tail(test_with_predictions_dt)
test_with_predictions_dt$Avg_prediction = rowMeans(test_with_predictions_dt[,2:101])
test_with_predictions_dt$Avg_prediction
tail(test_with_predictions_dt)
```
#Calculating Bias Square and Variance
```{r}
library(matrixStats)
test_with_predictions_dt$Bias_square = (test_with_predictions_dt$scaled_sound_pressure_level - test_with_predictions_dt$Avg_prediction)^2
DT_Regression_Bias_Square = mean(test_with_predictions_dt$Bias_square)
test_with_predictions_dt$Variance = rowVars(x = data.matrix(test_with_predictions_dt, rownames.force = NA), cols = c(-1,-102,-103))
tail(test_with_predictions_dt)
```

#Caculating the model Bias Square
```{r}
Bias_square_DT = mean(test_with_predictions_dt$Bias_square)
Bias_square_DT
```
#Calculating the model Variance
```{r}
Variance_DT = mean(test_with_predictions_dt$Variance)
Variance_DT
```

```{r}
Bias_square_DT + Variance_DT
```
# Calculating the Mean Squared Error
```{r}
calculationDT = test_with_predictions_dt[1:101]
calculationDT[2:101] = (calculationDT[2:101]-calculationDT[,1])^2
tail(calculationDT)
 
calculationDT$MSE = rowMeans(calculationDT[,2:101])
mean(calculationDT$MSE)
```

```{r}
calculationDT$MSE = rowMeans(calculationDT[,2:101])
mean(calculationDT$MSE)
```

----------------------------------------------Random_Forest--------------------------------------------------
# Bootstrap Sampling and Modeling(100samples, 100models)
```{r}
# parallel   ---------------------------------------------------------
library(foreach)
library(doParallel)
library(rpart)
library(randomForest)
#setup parallel back end to use 8 processors
cl<-makeCluster(6)
registerDoParallel(cl)

#  sample data 100 times 
length_divisor <- 1
predictions_rf<-foreach(m=1:100,.combine=cbind) %dopar% { 
        # using sample function without seed
  library(randomForest)   
  sampleRows <- sample(nrow(traindf), size=floor((nrow(traindf)/length_divisor)),replace = TRUE)
     fit_rf <- randomForest(scaled_sound_pressure_level ~ ., data = traindf[sampleRows,],keep.forest=TRUE, ntree=500)
     predictions_rf <- data.frame(predict(object=fit_rf, testdf[,-6]))
} 
stopCluster(cl)

library(pROC)
# Calculating the expected estimate
dim(predictions_rf)
test_with_predictions_rf = cbind(testdf,predictions_rf)
dim(test_with_predictions_rf)
tail(tail(test_with_predictions_rf))
test_with_predictions_rf = subset(test_with_predictions_rf[,c(-1,-2,-3,-4,-5)])
tail(test_with_predictions_rf)
test_with_predictions_rf$Avg_prediction = rowMeans(test_with_predictions_rf[,2:101])
test_with_predictions_rf$Avg_prediction
tail(test_with_predictions_rf)
```
# Calculating Bias Square and Variance
```{r}
library(matrixStats)
test_with_predictions_rf$Bias_square = (test_with_predictions_rf$scaled_sound_pressure_level - test_with_predictions_rf$Avg_prediction)^2
RF_Regression_Bias_Square = mean(test_with_predictions_rf$Bias_square)
test_with_predictions_rf$Variance = rowVars(x = data.matrix(test_with_predictions_rf, rownames.force = NA), cols = c(-1,-102,-103))
tail(test_with_predictions_rf)
```
```{r}
table(test_with_predictions_rf$predict.object...fit_rf..testdf....6...se.fit...TRUE...1...1)
```
# Caculating the model Bias Square
```{r}
Bias_square_RF = mean(test_with_predictions_rf$Bias_square)
Bias_square_RF
```
# Calculating the model Variance
```{r}
Variance_RF = mean(test_with_predictions_rf$Variance)
Variance_RF
```

```{r}
Bias_square_RF + Variance_RF
```
# Calculating the Mean Squared Error
```{r}
calculationRF = test_with_predictions_rf[1:101]
calculationRF[2:101] = (calculationRF[2:101]-calculationRF[,1])^2
tail(calculationRF)
 
calculationRF$MSE = rowMeans(calculationRF[,2:101])
mean(calculationRF$MSE)
```

```{r}
calculationRF$MSE = rowMeans(calculationRF[,2:101])
mean(calculationRF$MSE)
```

----------------------------------------------------SVM------------------------------------------------------
# Bootstrap Sampling and Modeling(100samples, 100models)
```{r}
# parallel   ---------------------------------------------------------
library(foreach)
library(doParallel)

#setup parallel back end to use 8 processors
cl<-makeCluster(2)
registerDoParallel(cl)

#  sample data 100 times 
length_divisor <- 1
predictions_svm<-foreach(m=1:100,.combine=cbind) %dopar% { 
        # using sample function without seed
  library(e1071)   
  sampleRows <- sample(nrow(traindf), size=floor((nrow(traindf)/length_divisor)),replace = TRUE)
     fit_svm <- svm(scaled_sound_pressure_level ~ ., data = traindf[sampleRows,],kernel = "linear")
     predictions_svm <- data.frame(predict(object=fit_svm, testdf[,-6]))
} 
stopCluster(cl)

library(pROC)
# Calculating the expected estimate
dim(predictions_svm)
test_with_predictions_svm = cbind(testdf,predictions_svm)
dim(test_with_predictions_svm)
tail(tail(test_with_predictions_svm))
test_with_predictions_svm = subset(test_with_predictions_svm[,c(-1,-2,-3,-4,-5)])
tail(test_with_predictions_svm)
test_with_predictions_svm$Avg_prediction = rowMeans(test_with_predictions_svm[,2:101])
test_with_predictions_svm$Avg_prediction
tail(test_with_predictions_svm)
```
# Calculating Bias Square and Variance
```{r}
library(matrixStats)
test_with_predictions_svm$Bias_square = (test_with_predictions_svm$scaled_sound_pressure_level - test_with_predictions_svm$Avg_prediction)^2
svm_Regression_Bias_Square = mean(test_with_predictions_svm$Bias_square)
test_with_predictions_svm$Variance = rowVars(x = data.matrix(test_with_predictions_svm, rownames.force = NA), cols = c(-1,-102,-103))
tail(test_with_predictions_svm)
```
# Caculating the model Bias Square
```{r}
Bias_square_svm = mean(test_with_predictions_svm$Bias_square)
Bias_square_svm
```
# Calculating the model Variance
```{r}
Variance_svm = mean(test_with_predictions_svm$Variance)
Variance_svm
```

```{r}
Bias_square_svm + Variance_svm
```
# Calculating the Mean Squared Error
```{r}

calculationsvm = test_with_predictions_svm[1:101]
calculationsvm[2:101] = (calculationsvm[2:101]-calculationsvm[,1])^2
tail(calculationsvm)
 
calculationsvm$MSE = rowMeans(calculationsvm[,2:101])
mean(calculationsvm$MSE)
```
